### 2015-08-23 manuscript/content/intro.tex
New code, data, and some paper changes.

Internet of Things (IoT) is a domain that represents the next most exciting technological revolution since the Internet. IoT will bring endless opportunities and impact every corner of our planet. In the healthcare domain, IoT promises to bring personalized health tracking and monitoring ever closer to the consumers. This phenomena is evidenced in a recent Wall Street Journal (June, 29, 2015) article entitled "Staying Connected is Crucial to Staying Healthy". Modern smartphones and related devices now contain more sensors than ever before. Data from sensors can be collected more easily and more accurately. In 2014, it is estimated that 46 million people are using IoT-based health and fitness applications. Currently, the predominant IoT-based health applications are in sports and fitness. However, disease management or preventive care health applications are becoming more prevalent. For example, the real-time preventive care applications such as those for detecting fall in elderly patients is one of the active research areas due to the aging population \cite{Tacconi:2011}. This paper describes a new category of real-time preventive health application that aims to reduce risky behaviors due to intoxication by predicting blood alcohol content (BAC) non-invasively and in real-time.

The Centers for Disease Control and Prevention (CDC) reported that between the years 2008 and 2010, roughly two-thirds of adults were drinkers, but adults between the ages 18 and 24 were most commonly associated with consumption of more than 5 drinks in one day \cite{Schoenborn:2013}. There is also a known causal relationship between alcohol consumption and risky behaviors \cite{Assaad:2006}. One of the most dangerous, worldwide (citation needed) problems of drinking is drunk driving. This problem is not only a hazard to the drunk drivers, but to pedestrians and other drivers. It is reported by the Bureau of Transportation Statistics that in 2010, 47.2\% of pedestrian fatalities and 39.9\% of vehicle occupant fatalities are caused by drunk driving \cite{Chambers:2012}.

Adults often find themselves on outings where they plan on drinking, but also on later driving home. At dangerous levels of intoxication, it can be difficult to judge ones own level of intoxication. Instead it would be better to get a definitive measurement; of the BAC, or simply a binary response, "drunk" or "not drunk." To do this, one can buy a breathalyzer and blow into it on every outing where drinking is involved. This is not a socially beneficial choice. The other option is to use a simple smartphone application to manually calculate BAC. However these demand a great deal of involvement from the user. To be practical, it would be useful to have some sort of non-invasive and accurate monitoring system that will warn its user if they become too intoxicated. A study conducted in Washington University shows that electronic intervention programs are more successful at reducing college student drinking than a general alcohol awareness program \cite{Ward:2015}. This system can also be used to warn friends and family, or prevent the operation of the user's car.

The main contributions of the paper are:

\begin{itemize}
	\item A machine learning model for the prediction of intoxication level from smartwatch sensor data.
	\item A general Android-based gateway system that can collect data from any type of physical or virtual sensor accessible by the host smartphone.
	\item An overall, sensor middleware system, including a geographic visualization service, that may be useful for answering many other important research questions.
\end{itemize}

>>>>

Internet of Things (IoT) is a domain that represents the next most exciting technological revolution since the Internet. IoT will bring endless opportunities and impact every corner of our planet. In the healthcare domain, IoT promises to bring personalized health tracking and monitoring ever closer to the consumers. This phenomena is discussed in a recent Wall Street Journal article, "Why Connected Medicine Is Becoming Vital to Health Care" \cite{Landro:2015}. Modern smartphones and smartwatches now contain a more diverse collection of sensors than ever before, and people are warming up to them. In January 2014, approximately 46 million US smartphone owners were reported to have used health and fitness applications \cite{Nielsen:2014}. Currently, sports and fitness are the predominant foci of IoT-based health applications. However, applications in disease management and health care are becoming increasingly prevalent. For example, detecting falling of elderly patients \cite{Tacconi:2011}.

Drunk driving is a dangerous, worldwide problem. This problem is not only a hazard to the drunk drivers, but also to pedestrians and other drivers. It is reported by the Bureau of Transportation Statistics that in 2010, 47.2\% of pedestrian fatalities and 39.9\% of vehicle occupant fatalities were caused by drunk driving \cite{Chambers:2012}. The Centers for Disease Control and Prevention (CDC) reported that between the years 2008 and 2010, roughly two-thirds of adults were drinkers, with adults between the ages 18 and 24 having the greatest association with heavy drinking \cite{Schoenborn:2013}.

At dangerous levels of intoxication, it can be difficult to judge ones own drunkenness. Instead it would be better to get a definitive measurement of the BAC, or simply a binary response: "drunk" or "not drunk." Compact breathalyzers are probably the best option at the moment, but these are not discreet and require deliberate action by the user. The other option is to use a smartphone application to manually calculate BAC, but these demand a greater deal of involvement from the user. To be practical, it would be useful to have some sort of non-invasive and accurate monitoring system that will warn its user if they become too intoxicated. It has been shown that electronic intervention programs are more successful at reducing college student drinking than a general alcohol awareness program \cite{Ward:2015}. This system can also be used to warn friends and family, or prevent the operation of the user's car.

The main contributions of our paper are: \begin{itemize}
	\item A machine learning model for the prediction of intoxication level from smartwatch sensor data.
	\item A general Android-based gateway system that can collect data from any type of physical or virtual sensor accessible by the host smartphone.
	\item An overall, sensor middleware system, including a geographic visualization service, that may be useful for answering many other important research questions.
\end{itemize}

### 2015-08-23 manuscript/content/related.tex
Switch format and other changes
New code, data, and some paper changes.

\section{Related Work}
%
%Include the current IoT middleware systems that you have surveyed. Should start this early.
%\input{related_work}

There is much research being done in the prediction or calculation of blood alcohol content.

The most common way to calculate BAC is through the Widmark equation. Posey and Mozayani in their paper \cite{Posey:2007} discuss the Widmark equation and improvements on its use by observing 5 previously published models and one new model for BAC prediction. The Widmark equation predicts BAC using elimination rates and the Widmark factor. Current use of the Widmark equation and BAC prediction is discouraged by experts due to inaccuracies, usually overestimation. Mathematical calculation of BAC is still sometimes used in DUI court cases, as there are not many alternatives, but it requires lab test results, information about what the person was drinking, how much and how long, BMI, and other hard to find details such as elimination rates. Some of this information has to be estimated or generalized which causes the calculation to be inaccurate. Also, due to the amount of information required, such a method is unusable by the general public for personal use. While BAC prediction is generally frowned upon in court cases, Posey and Mozayani make a point that using several different models provides a more accurate prediction than using the original Widmark equation alone.

Another known method of measuring BAC is through a transdermal ethanol sensor. Webster and Gabler \cite{Webster:2007} discussed the feasibility of transdermal ethanol sensors being integrated into ignition interlock systems in order to prevent drunk driving. A way to reduce the number of incidents related to drunk driving is to somehow stop people from driving drunk. Their approach to do this is to fit every vehicle with a ignition interlock system that responds to the driver's BAC. Current interlock systems like this involve a breathalyzer, but these are seen as expensive, bulky, and elicit a negative response from the general public. Transdermal ethanol sensing provides a more discreet way to measure BAC. However, according to the paper, an issue with transdermal ethanol sensing is that there is a significant delay, up to roughly an hour, from the time of alcohol consumption and detection at the skin.

>>>>

\section{Related Work}

There is much research being done in the detection, prediction, or calculation of a person's blood alcohol content. The most popular mathematical model seems to be the Widmark equation; as evidenced by its prevalence in internet search results. Posey and Mozayani discuss the Widmark equation and improvements on its use by observing 5 previously published models and one new model for BAC prediction (clarify!! point?) \cite{Posey:2007}. Because of the type of information required, such methods are not accurate for personal use by the general public.

Current interlock systems involve a breathalyzer, but these are seen as expensive, bulky, and are not are all discreet. Recently, there have been some investigations into using transdermal ethanol sensors to directly measure intoxication level. Webster and Gabler took a look at the feasibility of using transdermal ethanol sensors with ignition interlock systems in cars. Transdermal sensing provides a more discreet way to measure BAC, but this method is accompanied by the problem of a significant lag between transdermally sensed alcohol concentration and blood alcohol concentration. The lag is predictable, but not constant and requires additional information about the drinker \cite{Webster:2007}.
